---
title: "Statistical Rethinking - Chapter 4 Questions"
output:
  html_document:
    df_print: paged
---

```{r setup, include=F}
library(rethinking)
library(ggplot2)
data(Howell1)
d <- Howell1
mean_weight <- mean(d$weight)
sd_weight <- sd(d$weight)
d$weight.s <- (d$weight - mean_weight)/sd_weight
```

## Question 1
The weights listed below were recorded in the !Kung cencus, but heights were not recorded for these individuals. Provide predicted heights and 89% intervals (either HPDI or PI) for each of these individuals. That is, fill in the tble below, using model-based predictions.

```{r Q1}
set.seed(89)
# Data for which we need to predict
test_data <- data.frame(
  individual = c(1,2,3,4,5),
  weight = c(46.95, 43.72, 64.78, 32.59, 54.63)
  )

# Create model parameters
flist <- alist(
  height ~ dnorm(mu, sigma),
  mu <- a + b*weight, 
  a ~ dnorm(138, 100),
  b ~ dnorm(0, 10),
  sigma ~ dunif(0,50)
)

m4.1 <- map(flist, d)
post <- extract.samples(m4.1, 1000)

prediction <- 0

for (i in test_data$individual){
  #predicted_height_samples <- extract.samples(m4.1)$a + ((test_data[test_data$individual==i, ]$weight - mean_weight) * extract.samples(m4.1)$b / sd_weight)
  test_weight <- test_data[test_data$individual==i, ]$weight
  predicted_height_samples <- rnorm(10000 ,post$a + test_weight*post$b, post$sigma)
    
    
  prediction_temp <- data.frame(
    individual = i,
    mean_predicted_height = mean(predicted_height_samples), 
    lower_quantile_89 = PI(predicted_height_samples, 0.89)[1],
    upper_quantile_89 = PI(predicted_height_samples, 0.89)[2]
  )
  if(length(prediction)==1){
    prediction <- prediction_temp
  }
  else{
    prediction <- rbind(prediction, prediction_temp)
  }
}

test_data <- merge(test_data, prediction, by="individual")
test_data
```

## Question 2
Select out all the rows in the Howell1 data with ages below 18 years of age. If you do it right,
you should end up with a new data frame with 192 rows in it.

(a) Fit a linear regression to these data, using map. Present and interpret the estimates. For every
10 units of increase in weight, how much taller does the model predict a child gets?

```{r Q2a}
set.seed(89)
d_child <- d[d$age < 18,]

flist2 <- alist(
  height ~ dnorm(mu, sigma),
  mu <- a + b*weight, 
  a ~ dnorm(50, 20),
  b ~ dnorm(4, 10),
  sigma ~ dunif(0,50)
)

m4.2 <- map(flist2, d_child)
mean(extract.samples(m4.2,10)$a)
# A child with 0 weight would hypothetically be around 58 units tall
mean(extract.samples(m4.2,10)$b)
# For every extra 10 units of weight, the child would be around 27 units higher
```

(b) Plot the raw data, with height on the vertical axis and weight on the horizontal axis. Superimpose
the MAP regression line and 89% HPDI for the mean. Also superimpose the 89% HPDI for
predicted heights.

```{r Q2b}
weight.seq <- seq(from = 0, to = 50, by = 1)
mu <- link(m4.2, data=data.frame(weight = weight.seq))
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )
weight_modeled = data.frame(weight = weight.seq, mean = mu.mean, hdpi=t(mu.HPDI))
colnames(weight_modeled) <- c("weight","mean","lower","upper")

ggplot(weight_modeled, aes(x=weight, y=mean)) +
  geom_point() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha=0.5) +
  theme_bw() +
  geom_point(data=d_child, aes(x=weight, y=height), col="blue") +
  labs(x="weight", y="height")

```

(c) What aspects of the model fit concern you? Describe the kinds of assumptions you would
change, if any, to improve the model. You don’t have to write any new code. Just explain

```{r Q2c}
# The model is too simplistic for the data, the relationship between weight and height is not very linear (after a point, around 20kg). I would assume that the regression is logarithmic (height increases logarithmically with weight). 
```

## Question 3
Suppose a colleague of yours, who works on allometry, glances at the practice problems just
above. Your colleague exclaims, “That’s silly. Everyone knows that it’s only the logarithm of body
weight that scales with height!” Let’s take your colleague’s advice and see what happens.

(a) Model the relationship between height (cm) and the natural logarithm of weight (log-kg). Use
the entire Howell1 data frame, all 544 rows, adults and non-adults. Fit this model, using quadratic
approximation:
hi ~ Normal(mu(i); sigma)
mu(i) = a + b log(wi)
a ~ Normal(178; 100)
b ~ Normal(0; 100)
sigma ~ Uniform(0; 50)
where h(i) is the height of individual i and w(i) is the weight (in kg) of individual i. The function for
computing a natural log in R is just log. Can you interpret the resulting estimates?

```{r Q3a}
set.seed(89)

flist3 <- alist(
  height ~ dnorm(mu, sigma),
  mu <- a + b*log(weight), 
  a ~ dnorm(0, 100),
  b ~ dnorm(40, 40),
  sigma ~ dunif(0,50)
)

m4.3 <- map(flist3, d)
# When weight = 1 then the predicted height will be a. 
```

(b) Begin with this plot:
R code
plot( height ~ weight , data=Howell1 ,
col=col.alpha(rangi2,0.4) )
Then use samples from the quadratic approximate posterior of the model in (a) to superimpose on
the plot: 
(1) the predicted mean height as a function of weight, 

```{r Q3b1}
weight.seq2 <- seq(from = 1, to = 70, by = 1)
mu2 <- link(m4.3, data=data.frame(weight = weight.seq2))
mu.mean2 <- apply( mu2 , 2 , mean )
mu.HPDI2 <- apply( mu2 , 2 , HPDI , prob=0.97 )
weight_modeled2 = data.frame(weight = weight.seq2, mean = mu.mean2, hdpi=t(mu.HPDI2))
colnames(weight_modeled2) <- c("weight","mean","lower","upper")

# ppc
post <- extract.samples( m4.3 )
sim.height <- data.frame(weight = 1:70, t(apply(sim(m4.3, list(weight = weight.seq2)), 2, PI)))
colnames(sim.height) <- c ("weight", "sim_lower", "sim_upper")

weight_modeled2 <- merge(weight_modeled2, sim.height, by="weight")

ggplot(weight_modeled2, aes(x=weight, y=mean)) +
  geom_point() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.8, fill='red') +
  geom_ribbon(aes(ymin=sim_lower, ymax=sim_upper), alpha=0.6, fill="red") +
  geom_point(data=d, aes(x=weight, y=height), col="blue", alpha=0.3) +
  theme_bw() +
  xlim(5, 65) +
  ylim(40, 180)
```

(2) the 97% HPDI for the mean, and

```{r Q3b2}

```

(3) the 97% HPDI for predicted heights.

```{r Q3b3}

```

